Holomorphic field theories and higher algebra

### 2022-08-19 draft.tex
added reminder about how to truncate fact alg seciton
copied vacua stuff into document

@@

@@

>>>>

\owen{We can cut this sharply now. My main concerns are (1) to mention working on arbitrary complex manifolds and not just $\CC^d$, (2) to mention Beilinson-Drinfeld's approach, and (3) to explain chiralization/reduction for $d > 1$.}

@@

\owen{new version}

It is common in physics to study field theories defined on the spacetime $\RR^n$.
As the underlying spacetime $\RR^n$ is a group under addition (i.e., it is the translation group),
it is natural to ask for field theories with an action by translations.
That is, since the fields are sections of a fiber bundle over $RR^n$,
one can ask for an equivariant fiber bundle and ask for the equations of motion (or action functional) to be equivariant under translations.
In practice, physicists often work with field theories whose equations of motion are differential equations with constant coefficients.
(Nearly all the examples we've discussed in this survey satisfy this condition.)

Now such a field theory $\cT$ on $\RR^n$ has a space of solutions $\Sol_\cT(\RR^n)$ to its equations of motion.
A {\em classical vacuum} for the theory is a solution that this translation-invariant,
i.e., a constant solution.
As the solutions $\Sol(\RR^n)$ is a derived stack (at least if you adhere to the perspective we've advocated here), there is a derived stack of classical vacua
\[
\cM_{\rm vac}(\cT) = \Sol_\cT(\RR^n)^{\RR^n}
\]
given by the substack of constant solutions.
We call it the {\em moduli stack of classical vacua} for the theory~$\cT$.

Many stacks of interest to geometers and representation theorists appear naturally from holomorphic field theories.
For example, the holomorphic $\sigma$-model from $\CC^d$ into a target complex manifold $X$ has
\[
\cM_{\rm vac} \cong T^*[1-d]X.
\]
(If we consider the underlying underived stack, it is just $X$, unless $d = 1$.)
For supersymmetric gauge theories, the moduli stack of the holomorphic twist has the form
\[
\cM_{\rm vac} \cong V/G
\]
where $V$ is a (graded) $G$-representation and $V/G$ denotes the quotient stack.
The factor $V$ arises from solutions for the non-gauge field, which are sections of some associated vector bundle,
while the quotient by $G$ arises from the gauge transformation.
For another example, pure $\cN =1$ holomorphic Yang-Mills theory for group $G$ has \owen{true?}
\[
\cM_{\rm vac} \cong \fg/G,
\]
the adjoint quotient stack of the group.
For general supersymmetric field theories, which involve both a $\sigma$-model and gauge fields, the holomorphic twist has a moduli stack of classical vacua of the form
\[
\cM_{\rm vac} \cong X/G
\]
where $X$ is a derived stack with a $G$-action and $X/G$ denotes the quotient stack.

In studying questions about the classical vacua of supersymmetric field theories, however, physicists have not typically worked with the {\em stack}.
Instead, they have worked with a GIT quotient.
For instance, in the setting of a gauge theory with stack $V/G$ of classical vacua, it means they work with
\[
V//G = \Spec(\cO(V)^G),
\]
the scheme whose ring of functions are the $G$-invariant functions on the representation~$V$.
This affine scheme is known as the {\em affinization} of the stack $V/G$, as it is the best affine approximation to the stack (i.e., apply the left adjoint to the inclusion of affine schemes among stacks).
We will call it the moduli {\em space} of classical vacua and denote it by~$M_{rm vac}(\cT)$.

It is familiar to geometers that this affinization forgets interesting information about the stack because an invariant function must be constant on the closure of any orbit.
Thus, the canonical map
\[
V/G \to V//G
\]
often collapses orbits together.
An important example, pertinent to $\cN =1$ super Yang-Mills theory, is the adjoint quotient $\gl_N/ GL_N$, whose affinization is
\[
\gl_N//GL_N \cong \CC^N/S_N,
\]
by the Chevalley restriction theorem.
The canonical map here records the characteristic polynomial of a matrix and hence only knows its generalized eigenvalues and not its Jordan decomposition.
In other words, the canonical map collapses together Jordan blocks with the same diagonal entries.
For this reason, it loses details about the classical vacua remembered by the stack quotient.

\begin{rmk}
The physicists' focus on the moduli space is not unreasonable:
one must make measurements --- i.e., evaluate functions --- to identify the state of a physical system.
The moduli space captures what can be seen by measurements (notably here, the ``vacuum expectation values'' of different vacua).
On the other hand, the moduli stack knows each orbit, which provides a distinct solution to the equations of motion, whereas the space cannot distinguish between some vacua.
Moreover, the stack encodes much more data about the theory near each vacuum.
For instance, it remembers ``enhanced symmetries'' (such as the residual gauge fields) at each vacuum.
\end{rmk}

For the holomorphic QCD theory \owen{write data} the moduli stack of classical vacua is \owen{write it}
The moduli space is a well-known variety
\[
M_{rm vac}(\cT) \cong \Gr(N_f, N_c)
\]
by using classical invariant theory. \owen{show?}
A similar computation on the putative dual shows the following.

\begin{prop}
Seiberg dual theories have isomorphic moduli spaces.
\end{prop}

This relationship, which holds already for the supersymmetric theory before twisting, was one of the original pieces of evidence offered by Seiberg for his duality.

\owen{Sketch proof}: This is the first and second fundamental theorem of classical invariant theory!

\begin{rmk}
In the physics literature this isomorphism is given by explicitly matching gauge-invariant functions, which are often called mesons or baryons, as they can be identified with the data that identifies such particles.
These meson and baryons are precisely the usual generators of the ring of invariant functions, so an isomorphism of generators determines the desired isomorphism of rings. 
One must check that the relations are satisfied;
here it explains the superpotential on the magnetic side.
\end{rmk}

We want to point out, however, that {\em the moduli stacks are not isomorphic}.
A quick check is to note that there are orbits with different stabilizer groups.
Consider, for instance, the origin in each vector space.
On the electric side its stabilizer is $SL(N_c)$ and on the magnetic side it is $SL(N_f - N_c)$.
Thus each theory has a vacuum whose associated perturbative classical theory has no partner on the other side.
Hence Seiberg duality cannot hold at the stack level for the classical theories.

@@

******** Discussion about quantization **********

On the other hand, there is an open locus where the moduli stacks are equivalent:
the semi-stable points of $V/G$ \owen{fix description, as it's possibly just the stable locus}.
This locus determines the space $V//G$ and hence we have the isomorphism desired.
Let us try to unravel what this equivalence means for the field theories, both classical and quantum.

Consider any semistable point $x$ in $V/G$.
We view $x$ as a vacuum, which thus labels a perturbative classical theory describing the behavior of the theory $\cT$ near $x$ in the space of fields. 
For the vacuum $x$, the perturbative theory is equivalent to a holomorphic $\sigma$-model into the tangent space $T_x (V//G)$ \owen{Elaborate}. 
Hence the isomorphism of moduli spaces assures us there is a dual vacuum $x_D$ to $x$ in the dual theory, and its perturbative theory is equivalent to the perturbative theory for $x$
because there is an explicit identification of the tangent space to the corresponding point in the moduli space of the dual theory.
Thanks to this identification, a quantization of the theory labelled by $x$ determines a quantization of the theory labelled by $x_D$.
Hence on the semistable locus of $V/G$, we can guarantee an equivalence of quantizations.

This identification breaks down on the unstable locus, raising the following question.

{\bf Question:}
At the quantum level, what happens on the unstable locus?

In a related vein, we want to know what correspondence of quantum factorization algebras we expect from Seiberg duality. 
The mismatch for the moduli stack of classical vacua must be fixed somehow.

We note that in the physics literature, there are a number of modifications that are explored in this setting.
For instance, there are ``instanton corrections'' that cannot be found by perturbative techniques, although they can be treated in a mathematically precise way. 
\owen{Mention others? RG flow, symmetry breaking, ...}
It is interesting to wonder how these would appear in the holomorphic setting.

\owen{Other things: affinization as polydiscs algebras, comparison with mirror symmetry (note A-model involves instanton corrections)}

* Explain that you get a sheaf of polydiscs algebras over the moduli stack. We can view this as presenting a ``quantization of the moduli stack'' $\cM_{vac}^{\rm q}$ There is thus an ``affinization'' of the quantum moduli stack.

Conjecture: There is an equivalence of affinizations of the quantum moduli stacks for Seiberg dual theories.

This process probably looks away from the unstable locus in the same way that the classical part does.
